\begin{thebibliography}{99}

\bibitem{Jo2024}
Jo, S., Lee, G. M., \& Moon, I. (2024). Airline dynamic pricing with patient customers using deep exploration-based reinforcement learning. \emph{Engineering Applications of Artificial Intelligence}, 133, 108073.

\bibitem{Ahiska2015}
Ahiska, S.S., \& King, R.E. (2015). Inventory policy characterisation methodologies for a single–product recoverable manufacturing system. \emph{European Journal of Industrial Engineering}, 9(2), 222–243.

\bibitem{Ahmadi2014}
Ahmadi, M., \& Shavandi, H. (2014). Joint pricing and rationing in a production system with two demand classes. \emph{European Journal of Industrial Engineering}, 8(6), 836–860.

\bibitem{Aviv2008}
Aviv, Y., \& Pazgal, A. (2008). Optimal pricing of seasonal products in the presence of forward-looking consumers. \emph{Manufacturing \& Service Operations Management}, 10(3), 339–359.

\bibitem{Bautista2022}
Bautista-Montesano, R., Galluzzi, R., Ruan, K., Fu, Y., \& Di, X. (2022). Autonomous navigation at unsignalized intersections: A coupled reinforcement learning and model predictive control approach. \emph{Transportation Research Part C}, 139, 103662.

\bibitem{Bondoux2020}
Bondoux, N., Nguyen, A.Q., Fiig, T., \& Acuna-Agost, R. (2020). Reinforcement learning applied to airline revenue management. \emph{Journal of Revenue and Pricing Management}, 19(5), 332–348.

\bibitem{Cao2015}
Cao, P., Fan, M., \& Liu, K. (2015). Optimal dynamic pricing problem considering patient and impatient customers’ purchasing behaviour. \emph{International Journal of Production Research}, 53(22), 6719–6735.

\bibitem{Caro2012}
Caro, F., \& Gallien, J. (2012). Clearance pricing optimization for a fast-fashion retailer. \emph{Operations Research}, 60(6), 1404–1422.

\bibitem{DenBoer2015}
Den Boer, A.V. (2015). Dynamic pricing and learning: historical origins, current research, and new directions. \emph{Surveys in Operations Research and Management Science}, 20(1), 1–18.

\bibitem{Dixit2022}
Dixit, A., \& ElSheikh, A.H. (2022). Stochastic optimal well control in subsurface reservoirs using reinforcement learning. \emph{Engineering Applications of Artificial Intelligence}, 114, 105106.

\bibitem{Flapper2012}
Flapper, S., Gayon, J.-P., \& Vercraene, S. (2012). Control of a production–inventory system with returns under imperfect advance return information. \emph{European Journal of Operational Research}, 218(2), 392–400.

\bibitem{Gosavi2002}
Gosavi, A., Bandla, N., \& Das, T.K. (2002). A reinforcement learning approach to a single leg airline revenue management problem with multiple fare classes and overbooking. \emph{IIE Transactions}, 34(9), 729–742.

\bibitem{Hafez2019}
Hafez, M.B., Weber, C., Kerzel, M., \& Wermter, S. (2019). Efficient intrinsically motivated robotic grasping with learning-adaptive imagination in latent space. In \emph{2019 Joint IEEE 9th International Conference on Development and Learning and Epigenetic Robotics (ICDL-Epirob)}, IEEE, pp. 1–7.

\bibitem{LawheadGosavi2019}
Lawhead, R.J., \& Gosavi, A. (2019). A bounded actor–critic reinforcement learning algorithm applied to airline revenue management. \emph{Engineering Applications of Artificial Intelligence}, 82, 252–262.

\bibitem{Otero2015}
Otero, D.F., \& Akhavan-Tabatabaei, R. (2015). A stochastic dynamic pricing model for the multiclass problems in the airline industry. \emph{European Journal of Operational Research}, 242, 188–200.

\bibitem{Victor2020}
Victor, V. (2020). An experimental research on the consumer response towards online personalised pricing strategies: A comparative study between Indian and Malaysian online consumers. M.Sc. thesis, Szent István University, Gödöllő, Hungary.

\bibitem{Victor2018}
Victor, V., Thoppan, J., Jeyakumar Nathan, R., \& Farkas, M. (2018). Factors influencing consumer behavior and prospective purchase decisions in a dynamic pricing environment—An exploratory factor analysis approach. \emph{Social Sciences}, 7(9), 153.

\bibitem{Gao2022}
Gao, J., Le, M., \& Fang, Y. (2022). Dynamic air ticket pricing using reinforcement learning method. \emph{RAIRO-Operations Research}, 56, 2475–2493.

\bibitem{Lu2018}
Lu, R., Hong, S.H., \& Zhang, X. (2018). A dynamic pricing demand response algorithm for smart grid: Reinforcement learning approach. \emph{Applied Energy}, 220, 220–230.

\bibitem{Raju2006}
Raju, C.V.L., Narahari, Y., \& Ravikumar, K. (2006). Learning dynamic prices in electronic retail markets with customer segmentation. \emph{Annals of Operations Research}, 143, 59–75.

\bibitem{Rana2015}
Rana, R., \& Oliveira, F.S. (2015). Dynamic pricing policies for interdependent perishable products or services using reinforcement learning. \emph{Expert Systems with Applications}, 42, 426–436.

\bibitem{Collins2013}
Collins, L., \& Thomas, L. (2013). Learning competitive dynamic airline pricing under different customer models. \emph{Journal of Revenue and Pricing Management}, 12, 416–430.

\bibitem{Cao2020}
Cao, J., Liu, Z., \& Wu, Y. (2020). Learning dynamic pricing rules for flight tickets. In \emph{Knowledge Science, Engineering and Management}, Springer.

\bibitem{Liu2013}
Liu, Q., \& Zhang, D. (2013). Dynamic pricing competition with strategic customers under vertical product differentiation. \emph{Management Science}, 59, 84–101.

\bibitem{Levin2009}
Levin, Y., McGill, J., \& Nediak, M. (2009). Dynamic pricing in the presence of strategic consumers and oligopolistic competition. \emph{INFORMS Journal on Computing}, 55, 32–46.

\bibitem{Bertsaks2000}
Bertsaks, D. (2000). \emph{Dynamic Programming and Optimal Control}. Athena Scientific.

\bibitem{Bobbio2003}
Bobbio, A., Horváth, A., Scarpa, A., \& Telek, M. (2003). Acyclic discrete phase type distributions: Properties and a parameter estimation algorithm. \emph{Performance Evaluation}, 54(1), 1–32.

\bibitem{Skare2015}
Škare, V., \& Gospić, D. (2015). [Title not provided]. \emph{Vol. 63/No. 4}, 515–528. [Please update with correct title if known].

\bibitem{Fiig2010}
Fiig, T., Isler, K., Hopperstad, C., \& Belobaba, P. (2010). Optimization of mixed fare structures: Theory and applications. \emph{Journal of Revenue and Pricing Management}, 9, 152–170.

\bibitem{Williams2020}
Williams, K.R. (2020). Dynamic airline pricing and seat availability. \emph{Cowles Foundation for Research in Economics, Yale University}, Working Paper 208281.


\end{document}